Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-01-26 draft.tex
finish conclusion and abstract

    Abstract filler text blablabla!

@@

	challenge \cite{cuccovillo_open_2022}, one even promoted by regularatory bodies such as the 

@@

	the research discussed in the next section, I would like to use new combinations of features
	with a novel model architecture to generate explanations for fake or real audio
	classification, with the goal of singling out features that are useful not only for an
	accurate classification but also for an understandable explanation.

@@

@@

			standard size. The features used are as follows:

@@

					Harmonic-to-noise ratios: Inspired by previous research

@@

					Mel-spectrograms: A widely-used feature in synthetic audio detection, the
					mel-spectrogram is a spectrographic representation of an audio sample
					transformed in a manner to better represent human perception of frequencies
					\cite{qais_deepfake_2022}.

@@

					the mel-spectrogram, also widely-seen in research on synthetic audio detection.

@@

					Fundamental frequency lengths: The lengths of every fundamental frequency
					cycle in the sample. This has also been previously used in certain research on
					synthetic audio detection \cite{xue_audio_2022}. The output is one-dimensional
					with time as the axis.

@@

					Onset strength: As used in previous research \cite{li_comparative_2022}, the
					strengths of each onset in the audio sample, where an onset is point where
					there is a sudden rise in energy across the audio spectrum. This results in a
					one-dimensional output with time as the axis.

@@

					intensity is the total power at each point in the audio sample, given in db.
					This is calculated by creating a fourrier transformation of the sample and then
					summing across the frequency-axis for every point on the time-axis. Resulting
					in a one-dimensional output.

@@

					Pitch-fluctuations: Calculated for every sample in the audio as the difference

@@

					Shimmer features: Similarly defined to the jitter features, shimmer features
					instead use the amplitudes at each fundamental frequency period. Their purpose
					is to capture irregular vocal fold vibrations which may be an indication but
					not a guarantee of a synthetic voice \cite{chaiwongyen_deepfake-speech_2023}.

@@

		0.5. EER is approximately 0.03214. This outperforms other results tested in the original
		paper on this dataset \cite{muller_does_2022}. This result also exceeds some other results
		tested on this dataset \cite{yang_robust_2024}, produced using a variety of imperceptible
		features and trained on the ASV2019 dataset, while underperforming one other method
		\cite{ranjan_statnet_2022} that made use of raw waveform-based features, training on 70
		percent of the data in the dataset, validating on 10 percent and testing on the remaining
		20 percent.

@@

				Simple & until 10000 & 94.28\% & 0.04840

@@

@@

@@

@@

>>>>

	I present a method of classification of audio samples as synthetic or bona fide using a
	combination of features with a focus on aspects of a voice sample that can be heard by the
	human ear. I then use this classifier as a surrogate model to produce explanations using the
	LIME method, with the hope of introducing a kind of explainability into this area that has
	the potential to be understood by a wider range of people and not only people who are already
	familiar with the field. I also introduce two metrics in order to evaluate the usefulness of
	the features in the explanations, with the goal of making a conclusion about the potential
	usefulness of explanations generated using this classifier. Experiments were performed using
	the In-the-Wild dataset, a dataset created with a focus on testing generalization of previous
	methods of synthetic audio detection. The surrogate model demonstrated strong performance on
	the dataset, even in comparison to previous methods. However, the explanations generated by
	this method do not seem to offer potential for useful interpretation using the perceptible
	features, based on an evaluation of aggregated local explanations using the proposed metrics.
	I believe, however, that the proposed method offers a promising basis for future research into
	explanations using perceptible features.

@@

	challenge \cite{cuccovillo_open_2022}, one even promoted by regulatory bodies such as the

@@

	the research discussed in the next section, I would like to use new combinations of features,
	most of which are perceptible, with a novel model architecture to generate explanations for
	fake or real audio classification, with the goal of singling out features that are useful not
	only for an accurate classification but also for an explanation that can be understood by the
	average person, explained using features that can be heard by the human ear.

@@

			In order to increase the likelihood of the explanations being useful and
			understandable to the end user, a focus was placed on using multiple perceptible
			features as inputs for the classifier. Two imperceptible features were also added with
			the hypothesis, based on previous research, that they would positively of increasing
			the model performance.

@@

			standard size. A more detailed description of the individual features used follows:

@@

					Harmonic-to-noise ratios: A perceptible feature inspired by previous research

@@

					Mel-spectrograms: A widely-used imperceptible feature in synthetic audio
					detection, the mel-spectrogram is a spectrographic representation of an audio
					sample transformed in a manner to better represent human perception of
					frequencies \cite{qais_deepfake_2022}.

@@

					the mel-spectrogram also considered imperceptible, also widely-seen in
					research on synthetic audio detection.

@@

					Fundamental frequency lengths: Considered a perceptible feature, the
					fundamental frequency lengths (f0 lengths) are the lengths of every
					fundamental frequency cycle in the sample. This has also been previously used
					in certain research on synthetic audio detection \cite{xue_audio_2022}. The
					output is one-dimensional with time as the axis.

@@

					Onset strength: As used in previous research \cite{li_comparative_2022}, this
					perceptible feature represents the strengths of each onset in the audio
					sample, where an onset is point where there is a sudden rise in energy across
					the audio spectrum. This results in a one-dimensional output with time as the
					axis.

@@

					intensity, also a perceptible feature, is the total power at each point in the
					audio sample, given in db. This is calculated by creating a fourrier
					transformation of the sample and then summing across the frequency-axis for
					every point on the time-axis. Resulting in a one-dimensional output.

@@

					Pitch-fluctuations: Also classified as a perceptible feature, the pitch
					fluctuations are calculated for every sample in the audio as the difference

@@

					Shimmer features: Similarly defined to the jitter features, shimmer features,
					also considered perceptible, and used in the same research instead use the
					amplitudes at each fundamental frequency period. Their purpose is to capture
					irregular vocal fold vibrations which may be an indication but not a guarantee
					of a synthetic voice.
					\cite{chaiwongyen_deepfake-speech_2023}.

@@

@@

				Simple & until 10000 & 94.28\% & 0.04840 \\
				3 convolutional layers & until 10000 & 62.80\% & 0.37198

@@

		0.5, with an EER of approximately 0.03214. This outperforms other results tested in the
		original paper on this dataset \cite{muller_does_2022}. This result also exceeds some
		other results tested on this dataset, such as an example \cite{yang_robust_2024} produced
		using a variety of imperceptible features and trained on the ASV2019 dataset, while
		underperforming one other method \cite{ranjan_statnet_2022} that made use of raw
		waveform-based features, training on 70 percent of the data in the dataset, validating on
		10 percent and testing on the remaining 20 percent.

@@

			\begin{tabular}{c | c | c}
				Feature & importance & trust \\
				HNRs & 0.1140 & -0.0543 \\
				mel spectrogram & 0.1084 & -0.0378 \\
				MFCC & 0.7029 & 0.3753 \\
				f0 lengths & 0.0 & 0.0 \\
				onset strengths & 0.0 & 0.0 \\
				intensities & 0.0 & 0.0 \\
				pitch fluctuations & 0.0 & 0.0 \\
				jitter features & 0.1277 & 0.0250 \\
				shimmer features & 0.1227 & 0.0017
			\label{table:exp-results}
			\caption{Summary of the explanation results using both of the proposed features}
		In order to obtain a full picture of the quality of the explanations generated by this
		method, I have used the best performing model and generated explanations on the first 500
		samples of the testing data. I then summarized the explanations generated using both of
		the previously introduced metrics. A summary of the results of both metrics summarized by
		the feature is presented in table \ref{table:exp-results}. Measuring by both metrics, the
		MFCCs had the most positive and the most correct influence on the classification results of
		the model, With the jitter features as a distant second for trust and importance. This
		indicates that the MFCCs had not only the most influence, but also the most correct
		influence on the result of the classification, while the other features had either little,
		no or negative influence on the classification.

@@

	In this section, I will discuss the implications of the experimental results of the previous
	sections with a special focus being placed on the results of the explanation tests as it
	relates to the potential explainability for potential end users of such a system.
		\subsection{Viability and Potential of the Explanations}
		Because of this model's inclusion of several perceptible features, I made the hypothesis
		that this method has the potential to produce explanations that can be used as the basis
		for something more presentable to an end user. As can be seen in the previous section in
		table \ref{table:exp-results}, the perceptible features did not contribute as much to the
		end result of the classification as their imperceptible counterparts, in spite of their
		increased presence in the overall set of input features. Two features, the HNRs and the
		mel-spectrograms even had a net-negative impact on the correctness of the classifications
		in the experiments, leaving the question open, if future experiments may perform better
		without these features. Four perceptible features, f0 lengths, onset strengths,
		intensities and pitch fluctuations had an overall average influence on the classification
		of 0, measured on both metrics, indicating that the model did not learn a correlation
		between these features and the realness of audio samples in the dataset. In spite of this,
		I believe that it is still worth investigating if other kinds of perceptible features will
		have more of an influence on the overall classification or if there are other
		imperceptible features that could be used in the place of MFCCs that may provide a better
		balance between classification correctness and a proportional influence on the end result.
		\subsection{Limitations}
		In spite of the potential that I see with this method, it does also present some
		limitations when it comes to its real-world use, aside from the shortfalls of the
		perceptible features discussed in the previous section. The first such limitation is in
		the computational performance of the model. Because of its relative size, the model
		requires relatively capable hardware in addition to several gigabytes of memory in order
		to classify samples. This means that the devices this or a similar model can run on are
		limited, and the possibility of running this or a similar model locally on a handheld
		device such as a smartphone is limited to none at the time of writing. Another limitation
		with this method is that, even though it is known that perceptible features such as the
		ones used in my method are able to be heard by the human ear and are even used in
		medicine, for example, for diagnoses \cite{chaiwongyen_deepfake-speech_2023}, it is not
		yet clear to what extent exactly these features can be understood by the average person.
		Cursory research \cite{sharevski_blind_2024} is already present into what factors humans
		use to identify fake and real audio samples, as well as their performance, but the named
		aspects of the samples have not yet been mapped to such perceptible features as the ones
		used in this method, leaving the question of the usability of such explanations open.
		Finally, I have not addressed in my experiments how the features can be presented to the
		end user. Certain implementations of LIME already have means to present their results
		graphically, but the effectiveness of such methods has yet to be studied.

@@

	To conclude, I have presented a model architecture using a combination of various features,
	perceptible and imperceptible, with the hypothesis that such a model could be used to generate
	explanations of its own results using the LIME method that are more useful to a potential end
	user who is not already familiar with the field. This model architecture resulted in a high
	level of performance in comparison to several other methods, indicating the potential for
	consistent and correct classifications. The explanations, however, were not able to produce
	the desired result, placing most of the weight on MFCCs, with most perceptible features either
	having little, no, or negative impact on the end classification with the exception of jitter
	and shimmer.
	I believe that this is a method that offers a large potential, but is not useful for the
	generation of understandable explanations in its current state. I suggest, however, that this
	method could be changed and improved in the future with other features or other changes to
	base parameters that could potentially produce more useful explanations while either holding
	the accuracy high like it is now or even improving it further.
